---
title: "The making of cricket package yorkr - Part 3"
author: "Tinniam V Ganesh"
date: "March 16, 2016"
output: html_document
---

## Introduction
This is the 3rd part of my cricket package **yorkr** in R. In my 2 earlier posts  

a. [The making of cricket package yorkr - Part 1](https://gigadom.wordpress.com/2016/03/05/the-making-of-cricket-package-yorkr-part-1-2/). In this post I analyze the performance of team in a ODI match. The batting and bowling performances of the team were analyzed. This post also performed analyses of a country in all matches against another country for e.g. India vs All matches of Australia. The best performers with the bat and ball were determined, the best batting partenrships, the detailed performances of the bowlers of India and Australia in the confrontation was also analyzed.

b.[The making of cricket package yorkr - Part 2](https://gigadom.wordpress.com/2016/03/09/the-making-of-cricket-package-yorkr-part-2/)
This post includes all ODI matches between a country and others. For obvious reasons I have chosen India and selected all ODI matches played by India with other countries. This included batting and bowling performances of the country against all oppositions.

As mentioned in my earlier posts the data is taken from [Cricsheet](http://cricsheet.org/)

In this post I look at individual performances of batsmen and bowlers in ODIs. For this post I have chose Virat Kohli & Mahendra Singh Dhoni from India and AB Devilliers from South Africa. For bowling analyses I have selected RA Jadeja, Harbhajan Singh and Mitchell Johnson

```{r message=FALSE,warning=FALSE}
library(dplyr)
library(ggplot2)
library(yorkr)
library(reshape2)
library(gridExtra)
library(rpart.plot)
```



```{r cache=TRUE}
a <- getTeamBattingDetails("India",save=TRUE)
b <- getTeamBattingDetails("South Africa",save=TRUE)

```

Now get the details of the batsmen Virat Kohli and Mahendra Singh Dhoni from the saved India file and AB De Villiers from the saved South Africa file
```{r}
kohli <- getBatsmanDetails(team="India",name="Kohli")
dhoni <- getBatsmanDetails(team="India",name="Dhoni")
devilliers <-  getBatsmanDetails(team="South Africa",name="Villiers")

```

Now we can compare the performances of these 3 batsmen
```{r}

p1 <-batsmanRunsVsDeliveries(kohli,"Kohli")
p2 <- batsmanRunsVsDeliveries(dhoni, "Dhoni")
p3 <- batsmanRunsVsDeliveries(devilliers,"De Villiers")
grid.arrange(p1,p2,p3, ncol=3)

```

```{r}
kohli46 <- select(kohli,batsman,ballsPlayed,fours,sixes,runs)
p1 <- batsmanFoursSixes(kohli46,"Kohli")
dhoni46 <- select(dhoni,batsman,ballsPlayed,fours,sixes,runs)
p2 <- batsmanFoursSixes(dhoni46,"Dhoni")
devilliers46 <- select(devilliers,batsman,ballsPlayed,fours,sixes,runs)
p3 <- batsmanFoursSixes(devilliers46, "De Villiers")
grid.arrange(p1,p2,p3, ncol=3)

```

```{r}

p1 <-batsmanDismissals(kohli,"Kohli")
p2 <- batsmanDismissals(dhoni, "Dhoni")
p3 <- batsmanDismissals(devilliers, "De Villiers")
grid.arrange(p1,p2,p3, ncol=3)

```

```{r}

p1 <-batsmanMeanStrikeRate(kohli,"Kohli")
p2 <- batsmanMeanStrikeRate(dhoni, "Dhoni")
p3 <- batsmanMeanStrikeRate(devilliers, "De Villiers")
grid.arrange(p1,p2,p3, ncol=3)

```

```{r}
p1 <-batsmanMovingAverage(kohli,"Kohli")
p2 <- batsmanMovingAverage(dhoni, "Dhoni")
p3 <- batsmanMovingAverage(devilliers, "De Villiers")
grid.arrange(p1,p2,p3, ncol=3)
```

```{r}
batsmanRunsAgainstOpposition(kohli,"Kohli")
batsmanRunsAgainstOpposition(dhoni, "Dhoni")
batsmanRunsAgainstOpposition(devilliers, "De Villiers")

```

```{r}
batsmanRunsVenue(kohli,"Kohli")
batsmanRunsVenue(dhoni, "Dhoni")
batsmanRunsVenue(devilliers, "De Villiers")

```

```{r}
p1 <-batsmanRunsVsDeliveries(kohli,"Kohli")
p2 <- batsmanRunsVsDeliveries(dhoni, "Dhoni")
p3 <- batsmanRunsVsDeliveries(devilliers, "De Villiers")
grid.arrange(p1,p2,p3, ncol=3)
```

```{r}
par(mfrow=c(1,3))
par(mar=c(4,4,2,2))
batsmanRunsPredict(kohli,"Kohli")
batsmanRunsPredict(dhoni, "Dhoni")
batsmanRunsPredict(devilliers, "De Villiers")
dev.off()
```

```{r cache=TRUE}
c <- getTeamBowlingDetails("India",save=TRUE)
d <- getTeamBowlingDetails("Australia",save=TRUE)

```

jadeja <- getBowlerWicketDetails(team="India",name="Jadeja")
harbhajan <- getBowlerWicketDetails(team="India",name="Harbhajan")
ashwin <- getBowlerWicketDetails(team="India",name="Ashwin")
johnson <-  getBowlerWicketDetails(team="Australia",name="Johnson")


```{r}
p1 <- bowlerEconomyRate(jadeja,"RA Jadeja")
p2<-bowlerEconomyRate(harbhajan, "Harbhajan")
p3<-bowlerEconomyRate(ashwin, "Ashwin")
p4<-bowlerEconomyRate(johnson, "MG Johnson")
grid.arrange(p1,p2,p3,p4, ncol=2)
```

```{r}
p1<-bowlerMeanRuns(jadeja,"RA Jadeja")
p2<-bowlerMeanRuns(harbhajan, "Harbhajan")
p3<-bowlerMeanRuns(ashwin, "Ashwin")
p4<-bowlerMeanRuns(johnson, "MG Johnson")
grid.arrange(p1,p2,p3,p4, ncol=2)
```

```{r}
p1<-bowlerMovingAverage(jadeja,"RA Jadeja")
p2<-bowlerMovingAverage(harbhajan, "Harbhajan")
p3<-bowlerMovingAverage(ashwin, "Ashwin")
p4<-bowlerMovingAverage(johnson, "MG Johnson")
grid.arrange(p1,p2,p3,p4, ncol=2)
```

```{r}
p1<-bowlerWicketPlot(jadeja,"RA Jadeja")
p2<-bowlerWicketPlot(harbhajan, "Harbhajan")
p3<-bowlerWicketPlot(ashwin, "Ashwin")
p4<-bowlerWicketPlot(johnson, "MG Johnson")
grid.arrange(p1,p2,p3,p4, ncol=2)
```

```{r}
bowlerWicketsAgainstOpposition(jadeja,"RA Jadeja")
bowlerWicketsAgainstOpposition(harbhajan, "Harbhajan")
bowlerWicketsAgainstOpposition(ashwin, "Ashwin")
bowlerWicketsAgainstOpposition(johnson, "MG Johnson")

```

```{r}
bowlerWicketsVenue(jadeja,"RA Jadeja")
bowlerWicketsVenue(harbhajan, "Harbhajan")
bowlerWicketsVenue(ashwin, "Ashwin")
bowlerWicketsVenue(johnson, "MG Johnson")

```

```{r}
jadeja1 <- getDeliveryWickets(team="India",name="Jadeja",save=FALSE)
harbhajan1 <- getDeliveryWickets(team="India",name="Harbhajan",save=FALSE)
ashwin1 <- getDeliveryWickets(team="India",name="Ashwin",save=FALSE)
johnson1 <- getDeliveryWickets(team="Australia",name="MG Johnson",save=FALSE)
```

```{r}
par(mfrow=c(1,2))
par(mar=c(4,4,2,2))
bowlerWktsPredict(jadeja1,"RA Jadeja")
bowlerWktsPredict(harbhajan1,"Harbhajan Sigh")
dev.off()
```

```{r}
par(mfrow=c(1,2))
par(mar=c(4,4,2,2))
bowlerWktsPredict(ashwin1,"Ravichander Ashwin")
bowlerWktsPredict(johnson1,"MG Johnson")
dev.off()
```

knitr::kable(head(jadeja))